---
title: "shinyEvents Vignette"
author: "Sebastian Kranz"
date: "Sunday, January 11, 2015"
output: html_document
---

## Brief Summary

The shinyEvents package allows to write shiny applications that use classical event handlers, e.g. for button clicks, value changes, etc, instead of reactive programming. One does not write an explicit server function, but just adds event handlers to an `app` object. Widgets will be updated with explicit calls to setXXXX functions, like e.g. `setText(id, "New text")`. Widget values and event handlers can be set in a similar fashion for an app that has not yet started as for an already running app. 

## Installation

To install the package see the description on the package's Github page:

https://github.com/skranz/shinyEvents

## Motivation

RStudio's shiny is a great framework to generate web applications with R. In a classical shiny app, interactivity is not generated via event handlers but by `reactive programming`. For details, see the shiny documentation and tutorials under http://shiny.rstudio.com/.

While shiny's reactive programming model is great for smaller apps, I personally found it less useful for bigger applications that create a lot of interactive dynamic UI.

For example, when writing the initial shiny interface for my package RTutor (https://github.com/skranz/RTutor), I felt that some observers or render functions were triggered too frequently, and I was not sure where to best put the 'server code' of newly dynamically created objects. Of course, one can write large applications with reactivity, but given my limited understanding of the reactivity model, it just was hard for me...

Anyway, I generated the package `shinyEvents` to emulate the classical event-handling paradigm for shiny applications. At least, I find it useful for my own shiny projects. 


## Examples




```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
